---
title: "A tutorial on using the footyr package"
author: "Timothy Young"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{footyr_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(dplyr)
library(tibble)
library(jsonlite)
library(magick)
library(magrittr)
library(tidyr)
library(stringr)
library(httr)
```

## Welcome!
This page will teach you the basics and workings of using the `footyr` package in R. This package will make calls to the ISport API and in particular will call on the football (soccer) information from the ISport API. The API also has information for basketball, but this vignette will focus on football. Using the functions contained within this package, we can easily pull data about teams, players, leagues/cup competitions, statistics from around the world.

## About the ISport API
The [ISport API](https://www.isportsapi.com) provides data for more than 1600 football leagues and cups, and matches from over 110 countries.^[iSports football API, ProgrammableWeb, https://www.programmableweb.com/api/isports-football, accessed 26 Nov 2019] We are able to access live scores, player lineups, match schedules, player information and odds by utilising this API, and the company has been dedicated to sport for 12 years now.^[Ibid.]

The iSports API's security is through the use of a registered API key, which will be received upon sign-up. The output of the API is in JSON format. The API is mostly a paid API, but it allows for a 15 day free trial period, limited to 10,000 calls per day/endpoint, and a 2 min delay for live data. You may register for the API [here.](https://www.isportsapi.com/auth/register)

You may also find in-depth [documentation](https://www.isportsapi.com/docs.html) about the API on ISport's website.

## Overview of functions
footyr(0.1.0) contains 5 functions, all of which help to call and parse data from the API. The core functions all start with `my` to differentiate them.

* `ID()` to retrieve unique ID numbers for leagues and teams for further use in the package.
* `myleague()` retrieves data on football leagues/cups, including the league standing tables.
* `myplayer()` retreives data on squads, and player statistics.
* `myteam()` retrieves overview statistics on teams, past result and future fixture listings.
* `mytopstats()` retrieves league statistics of top players within the league

### Loading the package
Having installed the `footyr` package, let us load it.
```{r setup}
#Load the package 
library(footyr)

```


### Loading your API key
Having registered online and received your API key, you need to then store your API key in your REnviron. This is important as you will not be able to perform calls to the API without the key, and as this key is unique to you, you would need to keep it secret, much like a password.

**Important**: *You should store your API key under the name "ISPORT_KEY". This is crucial as this package relies on this.* 
```{r, eval=FALSE}
#Use this to set your API key, replacing "YOUR_API_KEY" with your actual key.
Sys.setenv("ISPORT_KEY" = "YOUR_API_KEY")

#Check your key
Sys.getenv("ISPORT_KEY")
```


### Retrieve ID numbers with `ID()`
This first function is basic, but is one of the most important in this package. It allows us to search leagues and teams by their names, and retrieve their ID numbers. These associated ID numbers are necessary, as the other functions in the package will make use of these ID numbers.

In this vignette we will be mostly using the English Premier League (EPL) and Arsenal as an example. Here we retrieve the ID numbers for them:
```{r}
ID(league_name = "English Premier League", team_name = "Arsenal")
```

We can also just retrieve the league ID number without specifying a team. Here with the Spanish La Liga:
```{r}
ID(league_name = "Spanish La Liga")
```


### Retrieve league data and standing tables with `myleague()`
This function that allows you to retrieve key information on football leagues/cups, including the league standing tables.

Let us retrieve information on the English Premier League (EPL), specifying the league ID code in the first argument (1639).
```{r}
myleague(leagueID = 1639)
```

Let us also view the current league table for the EPL with the `table` argument.
```{r}
myleague(leagueID = 1639, table = "no")
```


### Retrieve football team information and past/future fixtures with `myteam()`
This function that allows you to retrieve key statistics on a football team of your choice, including past results and future fixtures and club logo.

Let us then retrieve information on Arsenal FC, using the arguments `leagueID = 1639` (Remember that ID 1639 corresponds to the EPL) and `team_name = "Arsenal"`
```{r}
myteam(leagueID = 1639, team_name = "Arsenal")
```

Let us now choose to include the past 6 results for Arsenal, by specifying the `past_results` argument to "yes". We can see these results appended to the end of the dataframe.
```{r}
myteam(leagueID = 1639, team_name = "Arsenal", past_results = "yes")

```

Now, let us also look at the next 5 fixtures that Arsenal are due to play in the future. We make use of the `future` argument, and set it to 5. We can change how many future fixtures we want to view by changing the number, say `future = 10` if we want to view the next 10 fixtures for example.
```{r}
myteam(leagueID = 1639, team_name = "Arsenal", past_results = "yes", future = 5)
```


### Retrieve squad information of your football team with `myplayer()`
This function that allows you to retrieve squad statistics of a football team of your choice, including individual player and coaching staff statistics.

Let us once again retrieve squad information on the team Arsenal, who play in the EPL, specifying the team ID code (19) in the first argument `teamID`. We get the output in the form of a dataframe containing information of the entire squad in Arsenal.
```{r}
myplayer(teamID = 19)
```

We can also search for a particular player in the Arsenal team, by using the second argument `player_name`, input the player name as a string in this second argument. This provides a singular response corresponding to the player you choose and containing information about that player. Here we look for the player called "Mesut Ozil". You can also get a picture of the player in the Viewer tab of RStudio.
```{r}
myplayer(teamID = 19, player_name = "Mesut Ozil")
```


### Retrieve data and statistics of leagues' best players with `mytopstats()`
This function that allows you to retrieve complete statistics of the top players within a league, including additional stats not natively from the ISports API.

Looking once again at the EPL, we specify the `leagueID = 1639` and looking at players with the most goals, we specify `"Goals"` in the second argument `stat`.
```{r}
mytopstats(leagueID = 1639, stat = "Goals")
```

As an additional example, we can look at players in the Spanish La Liga, ranked by those with the most assists.
```{r}
mytopstats(leagueID = 1134, stat = "Assists" )
```

As an additional feature of this function, we can request for `more_details = "yes"` to include additional stats in the the stat chart. These stats are not present in the API data, but instead separately calculated using the function. 

Here we request top stats for the EPL by goals, and getting more stats like mins per goal, goal conversion rate and shot accuracy.
```{r}
mytopstats(leagueID = 1639, stat = "Goals", more_details = "yes")
```


## Summary
The functions in the `footyr` package should help you to better extract data from the ISport API more easily. While they are not exhaustive, they should provide sufficient data for you to work with. Remember that you may obtain data about any league/cup around the world! Good luck!

Note: If there are any problems/questions, do get in touch via my public Github account. Thank you.
